---
title: "s06 Exercises: Pivoting and tidy data"
author: "Ansley"
output: 
  html_document:
    keep_md: true
    theme: paper
---

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
lotr  <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/lotr_tidy.csv")
guest <- read_csv("https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/wedding/emails.csv")
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

# From Class:

## Pivot Longer

```{r}
#Making data untidy
haireye_untidy <- hair_eye %>% 
  mutate(eye = str_c(Eye, "_eyed")) %>% 
  pivot_wider(id_cols = Hair, names_from = eye, values_from = n)

haireye_untidy

#Making data tidy
haireye_untidy %>% 
  pivot_longer(cols = c(Blue_eyed, Brown_eyed, Green_eyed, Hazel_eyed),
               names_to = "eye",
               values_to = "n")

#An easier way to do this (especially with greater variables)
haireye_untidy %>% 
  pivot_longer(cols = contains("eyed"), #We know that all columns we want to turn into a single column have "eyed" in their names. 
               names_to  = "eye",
               values_to = "n")

#Or could do every column except for hair
haireye_untidy %>% 
  pivot_longer(cols = -Hair,
               names_to  = "eye",
               values_to = "n")

```

## Pivot Wider

```{r}

hair_eye %>% 
  pivot_wider(id_cols = Hair, #What are the "ID" columns
              names_from = Eye, 
              values_from = n)

```



## Exercise 1: Univariate Pivoting

Consider the Lord of the Rings data:

```{r}
lotr
```

1. Would you say this data is in tidy format?

*Yes, this data is tidy because each row is a single observation, each column is a unique variable, and each cell is a value.*

   <!-- Describe why or why not in this space. -->

2. Widen the data so that we see the words spoken by each race, by putting race as its own column.

```{r}
lotr_wide <- lotr %>% 
  pivot_wider(id_cols = c(-Race, -Words), 
              names_from = Race, 
              values_from = Words)
```

3. Re-lengthen the wide LOTR data from Question 2 above.

```{r}
lotr_wide %>% 
  pivot_longer(cols = c(Elf, Hobbit, Man), 
               names_to  = "Race", 
               values_to = "Words")
```


## Exercise 2: Multivariate Pivoting

Congratulations, you're getting married! In addition to the wedding, you've 
decided to hold two other events: a day-of brunch and a day-before round of 
golf. You've made a guestlist of attendance so far, along with food preference 
for the food events (wedding and brunch).

```{r}
guest %>% 
  DT::datatable(rownames = FALSE)
```

1. Put "meal" and "attendance" as their own columns, with the events living in a new column.

```{r}
guest_long <- guest %>% 
  pivot_longer(cols  = c(starts_with("attendance"), starts_with("meal")),
               names_to  = c(".value", "event"),
               names_sep = "_") #Column names must have a "separating" character

#So after the "_" our columns (cols =) all had the "event" (either wedding, brunch, or golf). That's why the name of event comes *second*, (names_to) because the name of the event is after the _ in our column names. This creates a column named "event" with these values in it.
#Before the _ our columns had Attendance or meal. So we left this as is. This created the two columns named "attendance" and "meal" with these values in them.
#I don't get the values_to though.
```

2. Use `tidyr::separate()` to split the name into two columns: "first" and 
"last". Then, re-unite them with `tidyr::unite()`.

```{r}
guest_long %>% 
  separate(name, into = c("First", "Last")) %>% 
  unite(col = "name", First:Last, sep = " ")
  # unite(col = "name", FILL_THIS_IN, sep = FILL_THIS_IN)

```

3. Which parties still have a "PENDING" status for all members and all events?

```{r}
guest_long %>% 
  group_by(party) %>% 
  summarize(all_pending = all(attendance == "PENDING"))
```

4. Which parties still have a "PENDING" status for all members for the wedding?

```{r}
guest_long %>% 
  group_by(party) %>% 
  filter(event == "wedding") %>% 
  summarize(pending_wedding = all(attendance == "PENDING"))
```


5. Put the data back to the way it was.

```{r}
guest_long %>% 
  pivot_wider(id_cols = c("party", "name"), 
              names_from  = event,
              values_from = c("attendance", "meal"))
```

6. You also have a list of emails for each party, in this worksheet under the 
   object `email`. Change this so that each person gets their own row. 
   Use `tidyr::separate_rows()`

```{r}
email %>% 
  separate_rows(guest, sep = ",")
```


## Exercise 3: Making tibbles

1. Create a tibble that has the following columns:

- A `label` column with `"Sample A"` in its entries.
- 100 random observations drawn from the N(0,1) distribution in the column `x`
  - "N" means the normal distribution. "(0, 1)" means mean = 0, sd = 1.
  - Use `rnorm()`
- `y` calculated as the `x` values + N(0,1) error. 

```{r}
n <- 100
tibble(label = "Sample A",
             x = rnorm(n),
             y = x + sd(x))
```


2. Generate a Normal sample of size 100 for each combination of the following 
means (`mu`) and standard deviations (`sd`).

```{r}
n <- 100
mu <- c(-5, 0, 5)
sd <- c(1, 3, 10)
nesting(mu = mu, sd = sd) %>% 
  group_by_all() %>% 
  mutate(z = list(rnorm(n, mu, sd))) %>% 
  unnest()
```


3. Fix the `experiment` tibble below (originally defined in the documentation 
of the `tidyr::expand()` function) so that all three repeats are displayed for 
each person, and the measurements are kept. Some code is given, but it doesn't
quite work. It needs a few adjustments. What are they?

```{r}
#Need to finish this one
experiment <- tibble(
  name = rep(c("Alex", "Robert", "Sam"), c(3, 2, 1)),
  trt  = rep(c("a", "b", "a"), c(3, 2, 1)),
  rep = c(1, 2, 3, 1, 2, 1),
  measurement_1 = runif(6),
  measurement_2 = runif(6)
)
experiment %>% expand(name, trt, rep)

?tibble()
```


