---
title: "Day3"
author: "Ansley"
date: "1/30/2020"
output:
  pdf_document: default
  html_document: default
---

---
title: "cm005 Worksheet: Exploring Geometric Objects"
output: 
  html_document:
    keep_md: true
    theme: paper
---

## Preliminary

Begin by loading the required packages. If you don't have these installed (or don't know whether you have them installed), you can install them by executing the following code in your console:


Now run this code chunk to load the packages:

```{r load packages, warning = FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(scales)) #lets you change scales (e.g. to dollars)
suppressPackageStartupMessages(library(tsibble))
knitr::opts_chunk$set(fig.align = "center")
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

## Exercise 1: Bar Chart Grammar (Together)

Consider the following plot. Don't concern yourself with the code at this point.

```{r, fig.width = 5, fig.height = 2}
gapminder %>% 
  filter(year == 2007) %>% 
  mutate(continent = fct_infreq(continent)) %>% #%>% means "then"
  ggplot(aes(continent)) +
  geom_bar() +
  theme_bw()
```


Fill in the seven grammar components for this plot.

| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `gapminder` |
| __aesthetic mapping__ | x = 'continent', y = 'count' |
| __geometric object__  | bars |
| scale                 | y: linear |
| statistical transform | none |
| coordinate system     | rectansgular |
| facetting             | none |


```{r}
#From class.
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha = 0.1) +
  scale_x_log10("GDP per capita", labels = scales::dollar_format()) +
  theme_bw() +
  ylab("Life Expectancy")
```

```{r}
#Building the scatterplot from scratch
ggplot(gapminder, aes(gdpPercap, lifeExp)) + #This alone will give you a blank graph. Need to add stuff. Only add stuff in aes function if you want it to map to a specific variable
  geom_point(mapping = aes(color = continent), alpha = 0.1) + #without alpha we can't distinguish 1 vs. 100 points in 1 spot
#alpha = transparency
#geom_point(mapping = aes(color = continent)) allows to distinguish continent by color, essentially adding another variable to the visualization!!
  scale_x_log10("GDP per capita", labels = scales::dollar_format()) + #dollar_format is a function, scales is a package, :: means look into the package scale for the function dollar_format
  theme_bw() +
  ylab("Life Expectancy") +
  scale_y_continuous(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90))
```


## Exercise 2: `ggplot2` Syntax (Your Turn)

The following is a tsibble (a special type of tibble containing time series data, which we'll see more of later), stored in the variable `mauna`, of CO$_2$ concentrations collected monthly at the Mauna Loa station.

Execute this code to store the data in `mauna`:

```{r}
(mauna <- tsibble::as_tsibble(co2) %>% 
   rename(month = index, conc = value))
```

```{r}
ggplot(mauna, aes(month, conc)) + 
 geom_line(col = "paleturquoise4") + 
  theme_bw() +
  xlab("Year") +
  ylab("C02")
```


### 2(a)

Produce a line chart showing the concentration over time. Specifically, the plot should have the following grammar components:

| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `mauna` |
| __aesthetic mapping__ | x: month, y: conc |
| __geometric object__  | lines |
| scale                 | linear |
| statistical transform | none |
| coordinate system     | rectangular |
| facetting             | none |

### 2(b)

It turns out that you're allowed to specify the aesthetic mappings in a `geom` layer instead of, or in addition to, in the `ggplot()` function, with the following rules:

- Aesthetics appearing in a `geom` layer apply only to that layer.
- If there are conflicting aesthetics in both the `ggplot()` function and the `geom` layer, the `geom` layer takes precedence.

The following code mistakenly puts the month variable on the y-axis. Fill in the `FILL_THIS_IN` so that you still obtain the same result as above.

```{r, fig.width = 5, fig.height = 2}
ggplot(mauna, aes(y = month)) +
  geom_line(aes(x = month, y = conc))+
  ylab("CO2")
```

### 2(c)

You can store the output of the plot in a variable, too. Store the plot from 2(a) in the variable named `p`, then add a layer to `p` that adds green points to the plot.

```{r, fig.width = 5, fig.height = 2}
p <- ggplot(mauna, aes(month, conc)) + 
 geom_line(col = "paleturquoise4") + 
  theme_bw() +
  xlab("Year") +
  ylab("C02")

p +
  geom_point(colour = "green")
```

### 2(d)

What's wrong with the following code? Fix it.

```{r, fig.width = 5, fig.height = 2}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha = 0.1)
```


### 2(e) BONUS

So you're a ggplot2 pro? Then, let's see this plot adapted to polar coordinates. Specifically:

- angle is month (January through December)
- radius is CO$_2$ concentration

The plot should look like a spiral, or concentric circles. 

```{r, fig.width = 5, fig.height = 2}

ggplot(mauna, aes(x = month, y = conc, fill = month)) +
  geom_line(col = 'purple', size = 2) +
  theme_bw() +
  coord_polar(theta = "x") +
  ylim(c(300, 400))
```

