---
title: "Functional Programming"
author: "Ansley"
date: "4/3/2020"
output: 
  html_document:
    keep_md: true
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

## Load Packages

```{r}
library(psychTools)
library(tidyverse)
library(gapminder)
library(broom)
library(dplyr)

```

```{r}
bfi <- psychTools::bfi
keys <- replace_na(psychTools::bfi.dictionary$Keying, 1)

bfi_mean <- bfi %>% mutate_at(names(bfi)[keys == -1], ~ 7 - .x) %>% 
  mutate(A = rowMeans(select(., A1:A5), na.rm = TRUE),
         C = rowMeans(select(., C1:C5), na.rm = TRUE),
         E = rowMeans(select(., E1:E5), na.rm = TRUE),
         N = rowMeans(select(., N1:N5), na.rm = TRUE),
         O = rowMeans(select(., O1:O5), na.rm = TRUE),
         gender = recode_factor(gender, `1` = "male", `2` = "female"),
         education = recode_factor(education, `1` = "some hs", 
                                              `2` = "hs", 
                                              `3` = "some college", 
                                              `4` = "college",
                                              `5` = "graduate degree")) %>% 
  select(gender:O)

```

## map() applies the requested function to each column or element

### “Take bfi_mean and select columns A through O. For each of these columns, predict the column using the lm() function with gender from the bfi_mean data frame as a predictor.”

```{r}
mods_gender <- 
  bfi_mean %>% 
  select(A:O) %>% 
  map(~ lm(.x ~ gender, data = bfi_mean))

```

## Produces a *list* of models (LM), one for each big 5 trait

```{r}
map(mods_gender, summary)
map(mods_gender, glance)
map(mods_gender, tidy)

```

```{r}
map(cars, mean)
```

## If you want to specify additional arguments, use ~

```{r}
map(cars, ~ mean(.x, trim = .1))
```

## When you use ~, in the function call, .x is a placeholder that means “whatever the current variable from the data frame is”

### So for above, that would be all the columns of car

```{r}
map_dbl() #returns numeric vector instead of list

rsq_gender <- 
  bfi_mean %>% 
    select(A:O) %>% 
    map_dbl(~ lm(.x ~ gender, data = bfi_mean) %>% 
                summary() %>% .$r.squared)
rsq_gender

# Gives us the R^2 of each.

```

```{r}
map_dfc() #combines into columns
map_dfr() #combines into rows

coef_gender <- 
  bfi_mean %>% 
    select(A:O) %>% 
    map_dfr(~ lm(.x ~ gender, data = bfi_mean) %>% 
                coef(),
            .id = "Trait")
coef_gender

```

```{r}
coef_summary_gender <- 
  bfi_mean %>% 
    select(A:O) %>% 
    map_dfr(~ lm(.x ~ gender, data = bfi_mean) %>% 
                tidy(conf.int = TRUE),
            .id = "Trait")
coef_summary_gender
```

### There are other map functions, see ?map()

## To prevent functions from getting too long...

```{r}
#Write function
get_coef_summary_gender <- function(.x, data) {
  mod <- lm(.x ~ gender, data = bfi_mean)
  return(tidy(mod, conf.int = TRUE))
}

#Run map()
coef_summary_gender <- 
  bfi_mean %>% 
    select(A:O) %>% 
    map_dfr(get_coef_summary_gender, .id = "Trait")
coef_summary_gender
```

## Advantages

### 1. Your code is often more readable. 
### 2. You can reuse the function multiple times throughout your analyses. 
### 3. Its easier to fix bugs or typos you might make because your code is only in one spot.

## Template

```{r}
function_name <- function(arguments) {
  # function code
  return(result) #optional
}
```

## Name
### How do we call it?

## Arguments
### What are its inputs?

## Function Code
### What do we do with inputs?

```{r}
# Can give default value with "="
# Example
plot_histogram <- function(data, color = "blue") {
  tbl <- tibble(data = data)
  ggplot(tbl, aes(x = data)) +
  geom_histogram(color = color, fill = color)
}

# If the user of this function doesn't specify, the histogram will be blue

simulated_data <- rnorm(100)
plot_histogram(simulated_data)

# Or you can specify 
plot_histogram(simulated_data, color = "black")

```

# 12.6 Activities

## Activity 1

### Do men and women differ on the Big Five traits? How big are the differences?

#### Yes, women are more agreeable (0.40), conscientious (0.19), and extraverted (0.24), and less neurotic (-0.32) and open (-0.10), than men.

```{r}
get_coef_summary_gender <- function(.x, data) {
  mod <- lm(.x ~ gender, data = bfi_mean)
  return(tidy(mod, conf.int = TRUE))
}

coef_summary_gender <- 
  bfi_mean %>% 
    select(A:O) %>% 
    map_dfr(get_coef_summary_gender, .id = "Trait")
coef_summary_gender

```

#### Plotting

```{r}
bfi_mean %>%
    ggplot(aes(gender, A, col = gender)) +
    stat_summary(fun.data = mean_cl_normal) +
    theme_bw() +
    xlab("Gender")+
    ylab("Agreeableness")

bfi_mean %>%
    ggplot(aes(gender, C, col = gender)) +
    stat_summary(fun.data = mean_cl_normal) +
    theme_bw() +
    xlab("Gender")+
    ylab("Conscientiousness")

bfi_mean %>%
    ggplot(aes(gender, E, col = gender)) +
    stat_summary(fun.data = mean_cl_normal) +
    theme_bw() +
    xlab("Gender")+
    ylab("Extraversion")

bfi_mean %>%
    ggplot(aes(gender, O, col = gender)) +
    stat_summary(fun.data = mean_cl_normal) +
    theme_bw() +
    xlab("Gender")+
    ylab("Openness")

bfi_mean %>%
    ggplot(aes(gender, N, col = gender)) +
    stat_summary(fun.data = mean_cl_normal) +
    theme_bw() +
    xlab("Gender")+
    ylab("Neuroticism")
```

### Do the Big Five traits increase or decrease with Age? Is there a linear or squared trend?

#### Yes, all increase with age.Also appears to be support for a squared trend for all, except neuroticism and openness (when comparing models).
```{r}

#Linear
get_summary_age_lin <- function(.x, data) {
  mod1 <- lm(.x ~ age, data = bfi_mean)
  return(tidy(mod1, conf.int = TRUE))
}

coef_summ_age_lin <- 
  bfi_mean %>% 
    select(A:O) %>% 
     map_dfr(get_summary_age_lin, .id = "Trait")
coef_summ_age_lin

# Squared
get_summary_age <- function(.x, data) {
  mod <- lm(.x ~ age + I(age^2), data = bfi_mean)
   return(tidy(mod, conf.int = TRUE))
}

coef_summ_age <- 
  bfi_mean %>% 
    select(A:O) %>% 
     map_dfr(get_summary_age, .id = "Trait")
coef_summ_age

# ANOVA
compare_models <- function(.x, data) {
  mod2 <- anova((lm(.x ~ age, data = bfi_mean)), (lm(.x ~ age + I(age^2), data = bfi_mean)))
  return(summary(mod2))
}

compare_models_out <- 
  bfi_mean %>% 
    select(A:O) %>% 
    map(compare_models)
compare_models_out


```

#### Graphing

```{r}
bfi_mean %>%
    ggplot(aes(age, A)) +
    stat_summary(fun.data = mean_cl_normal, geom = "point") +
    theme_bw() +
    xlab("Age")+
    ylab("Agreeableness")+
    geom_smooth(method = "lm") 

bfi_mean %>%
    ggplot(aes(age, C)) +
    stat_summary(fun.data = mean_cl_normal, geom = "point") +
    theme_bw() +
    xlab("Age")+
    ylab("Conscientiousness")+
    geom_smooth(method = "lm") 

bfi_mean %>%
    ggplot(aes(age, E)) +
    stat_summary(fun.data = mean_cl_normal, geom = "point") +
    theme_bw() +
    xlab("Age")+
    ylab("Extraversion")+
    geom_smooth(method = "lm") 

bfi_mean %>%
    ggplot(aes(age, O)) +
    stat_summary(fun.data = mean_cl_normal, geom = "point") +
    theme_bw() +
    xlab("Age")+
    ylab("Openness")+
    geom_smooth(method = "lm") 

bfi_mean %>%
    ggplot(aes(age, N)) +
    stat_summary(fun.data = mean_cl_normal, geom = "point") +
    theme_bw() +
    xlab("Age")+
    ylab("Neuroticism")+
    geom_smooth(method = "lm") 
```

### Do the Big Five traits differ across educational levels? Treat education as a categorical variable.

#### Yes

```{r}
get_summary_edu <- function(.x, data) {
  mod <- lm(.x ~ education, data = bfi_mean)
  return(summary(mod))
}

coef_summ_edu <- 
  bfi_mean %>% 
    select(A:O) %>% 
     map(get_summary_edu)
coef_summ_edu
```

#### Graphing

```{r}
bfi_mean %>%
    ggplot(aes(education, A, col = education)) +
    geom_point()+
    geom_boxplot()+
    theme_bw() +
    xlab("Education")+
    ylab("Agreeableness")

bfi_mean %>%
    ggplot(aes(education, C, col = education)) +
     geom_point()+
    geom_boxplot()+    
    theme_bw() +
    xlab("Education")+
    ylab("Conscientiousness")

bfi_mean %>%
    ggplot(aes(education, E, col = education)) +
    geom_point()+
    geom_boxplot()+
    theme_bw() +
    xlab("Education")+
    ylab("Extraversion")

bfi_mean %>%
    ggplot(aes(education, O, col = education)) +
    geom_point()+
    geom_boxplot()+    
    theme_bw() +
    xlab("Education")+
    ylab("Openness")
```

### How well do age and gender together predict the Big Five traits?

#### A: R2 = 0.07, C: R2 = 0.02, E: R2 = 0.01, N: R2 = 0.03, O: AR2 = 0.01


```{r}
rsq_age_gender <- 
  bfi_mean %>% 
    select(A:O) %>% 
    map_dbl(~ lm(.x ~ age + gender, data = bfi_mean) %>% 
                summary() %>% .$r.squared)
rsq_age_gender
```

### In your models in part (4), do the residuals appear to be normally distributed? Are they consistent across age ranges and gender groups?

#### Come back to this...

```{r}

```

